Self Regulation Ontology Retest Data Report

### 2017-03-17 SRO_Report.Rmd
working on new report

@@

@@

@@

clean_test_data = cbind(sub_id = test_data$sub_id, as.data.frame(apply(test_data[, -which(names(retest_data) %in% c("sub_id"))], 2, remove_outliers)))

@@

```{r}
#Remove outliers from retest data the way they have been from test data
clean_retest_data = cbind(sub_id = retest_data$sub_id, as.data.frame(apply(retest_data[, -which(names(retest_data) %in% c("sub_id"))], 2, remove_outliers)))

#Get list of transformed variables and the sign of their skew
logTr_vars = data.frame(grep('logTr', names(clean_test_data), value=T))
names(logTr_vars) = c('var')
logTr_vars$sign = ifelse(grepl(".ReflogTr",logTr_vars$var), 'negative', 'positive')
logTr_vars$var = gsub(".logTr|.ReflogTr", "", logTr_vars$var)

#Transform necessary vars and rename those columns
for(i in 1:length(names(clean_retest_data))){
  tmp_col = names(clean_retest_data)[i]
  if(tmp_col %in% logTr_vars$var){
    sign = logTr_vars$sign[which(tmp_col == logTr_vars$var)]
      if(sign == 'positive'){
        clean_retest_data[,tmp_col] = log(clean_retest_data[,tmp_col])
        names(clean_retest_data)[i] = paste0(tmp_col, '.logTr')
      }
      else if(sign == 'negative'){
        clean_retest_data[,tmp_col] = neg_log(clean_retest_data[,tmp_col])
        names(clean_retest_data)[i] = paste0(tmp_col, '.ReflogTr')
      }
    }
}
```

@@

Check transformations are correct
```{r}
retest_logTr_vars = data.frame(grep('logTr', names(clean_retest_data), value=T))
names(retest_logTr_vars) = c('var')
retest_logTr_vars$sign = ifelse(grepl(".ReflogTr",retest_logTr_vars$var), 'negative', 'positive')
retest_logTr_vars$var = gsub(".logTr|.ReflogTr", "", logTr_vars$var)
retest_logTr_vars == logTr_vars
```

@@

@@

#CHECK IF EVERYTHING IS ORDERED RIGHT
retest_subs_test_data$sub_id == retest_data$sub_id
```

@@

```{r}
all_columns <- unique(c(names(retest_data), names(retest_subs_test_data)))

matching_columns <- c()
for(i in 1:length(all_columns)){
  if(all_columns[i] %in% names(retest_data) & all_columns[i] %in% names(retest_subs_test_data)){
    matching_columns <- c(matching_columns, all_columns[i])
  }
}

matching_dv_columns <- c()

for(i in 1:length(matching_columns)){
  if(is.numeric(retest_data[,matching_columns[i]]) & is.numeric(retest_subs_test_data[,matching_columns[i]])){
    matching_dv_columns <- c(matching_dv_columns, matching_columns[i])
  }
}

cor_df <- data.frame(spearman = rep(NA, length(matching_dv_columns)))

row.names(cor_df) <- matching_dv_columns

for(i in 1:length(matching_dv_columns)){
  cor_df[matching_dv_columns[i], 'spearman'] <- cor(retest_data[,matching_dv_columns[i]],retest_subs_test_data[,matching_dv_columns[i]], method = 'spearman', use = "pairwise.complete.obs")
}
```

Distribution of all Pearson and Spearman correlations
```{r warning=FALSE, message=FALSE}
cor_df %>%
  ggplot(aes(spearman))+
  geom_histogram()+
  theme_bw()+
  ggtitle('Distribution of Spearman correlations for cleaned meaningful variables')
```

@@

## Distribution of each variable

Keeping a record of the distributions of the cleaned meaningful variables.

```{r}
# Function to check if the variables that went in to cor_df have normal distributions
diff_from_normal <- function(col){
  return(shapiro.test(col)$p.value < 0.05)
}

# Create and organize new df where distribution of vars data is stored. First for retest data only; then for test data for subjects who have retest data - restricting here since these are the subsets of data that went in to the correlation matrix. Merge and cleanup.

dist_retest_data <- as.data.frame(apply(retest_data[,matching_dv_columns], 2, diff_from_normal))
names(dist_retest_data) <- c('retest_not_normal')
dist_retest_data$dv <- row.names(dist_retest_data)
row.names(dist_retest_data) <- seq(1:nrow(dist_retest_data))
dist_retest_data = dist_retest_data %>% select(dv, retest_not_normal)

dist_retest_subs_test_data <- as.data.frame(apply(retest_subs_test_data[,matching_dv_columns], 2, diff_from_normal))
names(dist_retest_subs_test_data) <- c('test_not_normal')
dist_retest_subs_test_data$dv <- row.names(dist_retest_subs_test_data)
row.names(dist_retest_subs_test_data) <- seq(1:nrow(dist_retest_subs_test_data))
dist_retest_subs_test_data = dist_retest_subs_test_data %>% select(dv, test_not_normal)

dist_data <- merge(dist_retest_data, dist_retest_subs_test_data, by = 'dv')

rm(dist_retest_data, dist_retest_subs_test_data)

# Add whether dv's are tasks or survey to distribution dfs

dist_data = cor_df %>%
  select(dv, task) %>%
  left_join(dist_data, by = 'dv')
```

**What proportion is not normal?**  
A large proportion of the variables are not normal for data from both sources. Larger proportions of survey data are not normal compared to task data. 
```{r}
dist_data %>%
  group_by(task) %>%
  summarise(mean_retest_not_normal  = mean(retest_not_normal),
            mean_test_not_normal = mean(test_not_normal))
```

**Are the same vars that are not normally distributed in the retest data not normal in test data as well?**  
List of variables that have different distributions in the two time points.
```{r}
dist_data %>%
  filter(retest_not_normal != test_not_normal)
```

>>>>

```{r}
tst_mng_cln <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Complete_01-31-2017/meaningful_variables_clean.csv')

rtst_mng_cln <- read.csv('/Users/zeynepenkavi/Documents/PoldrackLabLocal/Self_Regulation_Ontology/Data/Retest_02-11-2017/meaningful_variables_clean.csv')

names(rtst_mng_cln)[which(names(rtst_mng_cln) %in% names(tst_mng_cln) == FALSE)]

names(tst_mng_cln)[which(names(tst_mng_cln) %in% names(rtst_mng_cln) == FALSE)]
```

@@

  cat(rep('*', 40))
  cat('\n')
  cat(paste0(length(names(positive_subset)) ,' data positively skewed data were transformed:'))
  cat('\n')
  cat(names(positive_subset), sep = '\n')
  cat(rep('*', 40))
  cat('\n')

@@

  cat(rep('*', 40))
  cat('\n')
  cat(paste0(length(names(negative_subset)) ,' data negatively skewed data were transformed:'))
  cat('\n')
  cat(names(negative_subset), sep = '\n')
  cat(rep('*', 40))
  cat('\n')

@@

clean_test_data = cbind(sub_id = test_data$sub_id, as.data.frame(apply(test_data[, -which(names(test_data) %in% c("sub_id"))], 2, remove_outliers)))

@@

Remove outliers from retest data the way they have been from test data
```{r}
clean_retest_data = cbind(sub_id = retest_data$sub_id, as.data.frame(apply(retest_data[, -which(names(retest_data) %in% c("sub_id"))], 2, remove_outliers)))
```

Get list of transformed variables and the sign of their skew
```{r}
logTr_vars = data.frame(grep('logTr', names(clean_test_data), value=T))
names(logTr_vars) = c('var')
logTr_vars$sign = ifelse(grepl(".ReflogTr",logTr_vars$var), 'negative', 'positive')
logTr_vars$var = gsub(".logTr|.ReflogTr", "", logTr_vars$var)
```

Transform the t2 columns the same way t1 columns were transformed.

```{r}
for(i in 1:length(names(clean_retest_data))){
  tmp_col = names(clean_retest_data)[i]
  if(tmp_col %in% logTr_vars$var){
    sign = logTr_vars$sign[which(tmp_col == logTr_vars$var)]
      if(sign == 'positive'){
        clean_retest_data[,tmp_col] = log(clean_retest_data[,tmp_col])
        names(clean_retest_data)[which(names(clean_retest_data) == tmp_col)] = paste0(tmp_col, '.logTr')
        cat(paste0(tmp_col ,' was positively skewed and transformed.'))
        cat('\n')
      }
      else if(sign == 'negative'){
        clean_retest_data[,tmp_col] = neg_log(clean_retest_data[,tmp_col])
        names(clean_retest_data)[which(names(clean_retest_data) == tmp_col)] = paste0(tmp_col, '.ReflogTr')
        cat(paste0(tmp_col ,' was negatively skewed and transformed.'))
        cat('\n')
      }
    }
}
```

@@

Check transformations are correct
```{r}
retest_logTr_vars = data.frame(grep('logTr', names(clean_retest_data), value=T))
names(retest_logTr_vars) = c('var')
retest_logTr_vars$sign = ifelse(grepl(".ReflogTr",retest_logTr_vars$var), 'negative', 'positive')
retest_logTr_vars$var = gsub(".logTr|.ReflogTr", "", retest_logTr_vars$var)
retest_logTr_vars = retest_logTr_vars %>% arrange(var)
logTr_vars = logTr_vars %>% arrange(var)
retest_logTr_vars == logTr_vars
```

@@

Switch df names

```{r}
raw_retest_data = retest_data
raw_test_data = test_data
test_data = clean_test_data
retest_data = clean_retest_data
rm(clean_test_data, clean_retest_data)
```

@@

#CHECK IF EVERYTHING IS ORDERED RIGHT
as.character(retest_subs_test_data$sub_id) == as.character(retest_data$sub_id)
```

@@

cor_df <- data.frame(spearman = rep(NA, length(numeric_cols)))

row.names(cor_df) <- numeric_cols

for(i in 1:length(numeric_cols)){
  cor_df[numeric_cols[i], 'spearman'] <- cor(retest_data[,numeric_cols[i]],retest_subs_test_data[,numeric_cols[i]], method = 'spearman', use = "pairwise.complete.obs")
}
```

Distribution of Spearman correlations
```{r warning=FALSE, message=FALSE}
cor_df %>%
  ggplot(aes(spearman))+
  geom_histogram()+
  theme_bw()+
  ggtitle('Distribution of Spearman correlations for cleaned meaningful variables')
```

@@

## ICC for all cleaned meaningful variables

```{r}
library(psych)
sf <- matrix(c(9,    2,   5,    8,
6,    1,   3,    2,
8,    4,   6,    8,
7,    1,   2,    6,
10,   5,   6,    9,
6,   2,   4,    7),ncol=4,byrow=TRUE)
colnames(sf) <- paste("J",1:4,sep="")
rownames(sf) <- paste("S",1:6,sep="")
sf
ICC(sf)
```

```{r}
library('irr')

time1 <- c(0:10)
time2 <- c(1:11)
times <- cbind(time1,time2)
rm(time1, time2)
times
```

```{r}
icc(times,model = "twoway",
    type = "consistency",
    unit = "single",
    r0 = 0,
    conf.level = 0.95)
```

```{r}
#>>ICC=1 for consistency

icc(times,model = "twoway",
    type = "agreement",
    unit = "single",
    r0 = 0,
    conf.level = 0.95)
```

```{r}
tmp = data.frame(persons=c(1,2,3,4,5,6),
                 rater1=c(9,6,8,7,10,6),
                 rater2=c(2,1,4,1,5,2),
                 rater3=c(5,3,6,2,6,4),
                 rater4=c(8,2,8,6,9,7))

ICC(tmp%>%select(-persons))

tmp = tmp %>% gather(rater,score,-persons)
tmp
summary(aov(score ~ rater+Error(factor(persons)), tmp))
```

```{r}
match_t1_t2 <- function(dv_var, t1_df = test_data, t2_df = retest_data, merge_var = 'sub_id'){
  df = merge(t1_df[,c(merge_var, dv_var)], t2_df[,c(merge_var, dv_var)], by = merge_var)
  df = df %>% 
    gather(time, score, -sub_id) %>%
    separate(time, c("dv","drop","time"), sep="[.]") %>%
    select(-drop)%>%
    mutate(time=ifelse(time=="x",1,2))
  return(df)
}

tmp = match_t1_t2("adaptive_n_back.mean_load")

summary(lmer(score ~ (1|sub_id)+factor(time),data=tmp))

```

```{r}
summary(aov(score ~ factor(time) + Error(sub_id),tmp))
```

----------------------------------------------------------------------------------------------------------------

```{r}
icc_df <- data.frame(icc = rep(NA, length(numeric_cols)))

row.names(icc_df) <- numeric_cols

for(i in 1:length(numeric_cols)){
  icc_df[numeric_cols[i], '...'] <- ...
}
```

----------------------------------------------------------------------------------------------------------------
